---
title: "Final for Env S 193DS"
format: html
---

# Set Up

```{r}
library(tidyverse) #loading in tidyverse
library(here) #loading in here
library(janitor) #loading in janitor
library(readxl) #loading in readxl
library(ggeffects) #loading in ggeffects
library(knitr) #loading in knitr
library(MuMIn) #loading in MuMIn
kelp <- read_csv(
  here(
    "Data", 
    "kelp.csv"
  ))
mopl <- read_csv(
  here(
    "Data", 
    "MOPL_nest-site_selection_Duchardt_et_al._2019.csv"
  )
)
```

# Problem 1. San Joaquin River Delta

## a. We determined that there is not enough evidence to suggest a difference between the median flooded wetland area across water year classification (wet, above normal, below normal, dry, critical drought).

## b. A boxplot could be created to highlight the medians and interquartile ranges of the median flooded wetland area across the different water year classifications. On the x-axis would be the different water year classifications and on the y-axis would be the recorded flooded wetland area, including the medians. In addition to the boxplot, the points could be jittered vertically to further show the data.

## c.

## d.

# Problem 2. Data Visualization

## a.

```{r}
kelp_clean <- kelp |> 
  # function 3
clean_names() |>  #removes underscores and capital letters
  
  # function 6
filter(site %in% c("IVEE", "NAPL", "CARP")) |>  #filtering by three sites
  
  # function 8
mutate(site = case_when( #changing site names
  site == "NAPL" ~ "Naples", #changing site NAPL to Naples
  site == "IVEE" ~ "Isla Vista", #changing site IVEE to Isla Vista
  site == "CARP" ~ "Carpinteria")) |> #changing site CARP to Capinteria 

  # function 2
mutate(site = as_factor(site), #converts the site name to a categorical variable
       site = fct_relevel(site,
                               "Naples", "Isla Vista", "Carpinteria")) |> #changes the order of factor levels
  # function 1
filter(!(date == "2000-09-22" & transect == "6" & quad == "40"))|> #filtering by the date, transect, and quad number
    # function 5
group_by(site, year) |> #grouping by site name and year
    # function 7
summarize(mean_fronds = mean(fronds)) |>  #creating a column for the mean 
    # function 4
ungroup()   #ungrouping any prior groupings

```

## b.

## c.

```{r}
ggplot(data = kelp_clean, #using kelp_clean df
       aes(x = year, #year on x axis
           y = mean_fronds, #mean_fronds on y axis
           color = site)) + #grouping by site
  geom_point() + #inserting points on graph
  geom_line() + #adding line to the graph
  labs(x = "Mean kelp fronds per year", #titling the x axis
       y = "Year", #titling the y axis
       title = "Sites vary in mean kelp fronds per year") + #adding a title
  scale_color_manual(values = c("Naples" = "cyan4", #changing the color of Naples points and line
                                "Isla Vista" = "burlywood3", #changing the color of Isla Vista points and line
                                "Carpinteria" = "grey")) + #changing the color of Carpinteria points and line
  theme_minimal() + #changing theme to minimal
  theme(plot.title = element_text(hjust = 0.6), #altering plot title position to be centered
        text = element_text(family = "Futura")) #changing the overall text font 
```

## d.

Isla Vista had the most variable mean kelp fronds per year because it has the highest peaks overall and the greatest difference in increases and decreases per year (represented by the changes in slope of line between each yearly point). Carpinteria had the least variable mean kelp fronds per year because the line has fairly consistent minimums and maximums (represented by the highest and lowest points).

# Problem 3. Data Analysis

## a.

The 1s and 0s represent whether the site was used for nesting by Mountain Plovers, where 0 is the absence and 1 is the presence.

## b.

The variable for distance to prairie dog colony edge is a continuous variable. This could influence the probability of Mountain Plover nest site use because it increases the vulnerability of the nesting site due to altered vegetation composition, which is included in the methods and discussion section of the paper. The variable for visual obstruction is a continuous variable. It could influence the probability of Mountain Plover nest site use because it can impact the visibility of the nest to predators, which is included in discussion section of the paper.

## c.

| Model number | Distance to prairie dog colony edge | Visual obstruction | Predictors |
|-------------|:-----------:|:-----------:|:--------------------:|
| 0 |  |  | no predictors (null model) |
| 1 | X | X | all predictors (full model) |
| 2 |  | X | only visual obstruction |
| 3 | X |  | only distance to prairie dog colony edge |

## d.
```{r}
#| message: false
#| warning: false
#| include: false
mopl_clean <- mopl |> 
  clean_names() |> 
  mutate(used_bin = case_when(
    id == "used" ~ 1,
    id == "available" ~ 0)) |> 
  select(used_bin, edgedist, vo_nest)
  
# model 0: null model
model0 <- glm(used_bin ~ 1, data = mopl_clean)

# model 1: all predictors
model1 <- glm(used_bin ~ vo_nest + edgedist, data = mopl_clean)

# model 2: only visual obstruction
model2 <- glm(used_bin ~ vo_nest, data = mopl_clean)

# model 3: only distance to prairie dog colony edge
model3 <- glm(used_bin ~ edgedist, data = mopl_clean)
```

## e.
```{r}
AICc(
  model0,
  model1, 
  model2, 
  model3
) |> 
  arrange(AICc)
```

The best model as determined by Akaike's Information Criterion (AIC) was the model with only visual obstruction. 

## f.

## g.

## h.

## i.

## j.

# Problem 4. Affective and Exploratory Visualizations

## a.

The visualizations from Homework 2 and Homework 3 are similar in the overall outline of the visualization. They have the same x and y axes and have similar shapes, given that the visualization for Homework 3 is an abstract version of Homework 2. However, the final product for the affective visualization differs in that it displays another variable than the initial visualization it was based on does. The affective visualization combines the two initial visualizations in Homework 2, as it displays the school and non school days, as well as a time series plot. In these visualizations, I see a trend in my wake up times on weekends being later than on week days. This is consistent between all visualizations, as they all display, in some way, wake up times for weekends and school days. In week 9, the primary feedback I got was to add in another variable to my visualization. One idea presented was to add in the time of my first obligation as clouds or something similar. I realized I wouldn't be able to do this because there was such a big contrast between my first obligation and my wake up time. Instead, I added another color to represent when the weekends were to differentiate and highlight the differences in wake up times, which was also my primary predictor variable.

## b.

## c.

## d.

## e.

## f.

## g.

I completed the activity in section.
